Processing of visual and non-visual naturalistic spatial information in the "parahippocampal place area"

### 2021-02-07 paper/main.tex
adjusting ending of discussion

% previous studies
Previous studies reported increased hemodynamic activity in the PPA when
participants were watching static pictures of landscape compared to pictures of
faces or objects (e.g. \citep{epstein1998ppa, epstein1999parahippocampal})
whereas one study using spoken sentences reported mixed results
\citep{aziz2008modulation}.
% current study: similar
In line with previous studies, we also operationalized spatial perception via
general linear model (GLM) $t$-contrasts.
% current study: dissimilar
Contrary to previous studies, our current approach is not based on carefully
designed sets of stimuli but based on annotated stimulus features incidentially
occurring in two naturalistic stimuli.
% group AV
On a group level, whole brain-analysis results for the movie stimulus show
signifcantly increased hemodynamic activity spatially overlapping with a
traditionally localized PPA but also extending into earlier visual cortices.
% group AD
Results for the audio-description reveal significant activation in the PPA but
restricted to its anterior part.
% individual level
On an individual level, results of the audio-description show significant
bilateral clusters in the anterior PPA of nine subjects and significant
unilateral clusters in one subjects.
% meaning 1
Results suggest that increased activation in the PPA during the perception of
static pictures generalizes to the perception of spatial information embedded in
a movie and an exclusively auditory stimulus.
% meaning 2
Our results provide further evidence that the PPA can be divided into functional
subregions. An auditory narrative might substitute a visual paradigm to localize
the PPAs anterior part in visually impaired subjects.

@@

% intro
First, we analyzed data from the movie stimulus offering ecologically more valid
but still visual stimulation.
%
An early study \citep{bartels2004mapping} that manually annotated the content
(color, faces, language, and human bodies) of movie frames found that
specialization of functional areas is preserved during movie watching.
%
Contrary to \citep{bartels2004mapping} who annotated the movie frames' content
(color, faces, language and human bodies), we categorized each frame after a cut
independent of its content. \todo{Bartels' clusters are actually pretty big too}
% hypo
Despite this ad hoc approach, we hypothesized that [confounding variables
could/would average out and] our whole-brain analyses would lead to significant
clusters in the PPA.
% results
Group results of the AV stimulus' primary contrast yielded a cluster that span
the group overlap of individual PPA ROIs from anterior to posterior.
% difference in location/size to AD and VIS
The cluster also extends into medial regions (dorsal precuneus and dorsal
posterior cingulate cortices), and into more posterior regions (occipital
fusiform gyrus and intracalcarine cortex) bilaterally.
% confound of higher perception
The larger extension into more posterior, primary visual areas can be attributed
to the averaging process that did not control for confounding visual features or
the attentional focus of study participants.
%
\todo[inline]{I am pretty surprised the clusters are not all over the brain.  I
do not know where/how do phrase it but should be mentioned imo}
% solution
Future studies that aim to use a movie to localize visual areas should
extensively annotate the content of frames (e.g. using the open-source solution
``Pliers'' for feature extraction from a visual naturalistic stimulus;
\citep{mcnamara2017developing}) and record the subject's visual attention via
eye-tracking.

% exploited cinematographic tendencies
We here exploited a cinematographic confound in the stimulus structure (cf.
\ref{rationale}).
% establishing shots vs. e.g. over-the-shoulder shots
Film directors tend to establish the geometry of locations earlier in the movie
and tend to later focus more on persons and objects relevant to the evolved
\citep{brown2012cinematography, mercado2011filmmakers}.
%
\todo[inline]{nice that exploited confound structure is mentioned but what to do
with it? hey, look, it works? movies are nice but have idiosyncrasies? so, maybe
you should be aware of these idiosyncrasies (and control for it) in case you
wanna use movies for research?}

% summary: nevertheless
\todo[inline]{Overall summary for AV \& concluding statement!?!}

\todo[inline]{so, yes, what is the main message of this
``cross-modal-control-movie-thing''??}

@@

% intro
Second, we analyzed data from the audio-description.
% hypothesis
Encouraged by \citep{aziz2008modulation}, we also hypothesized that spatial
information embedded in an exclusively auditory stimulus would correlate with
increased hemodynamic activity in the PPA.
% diff to Aziz (2008)
\citep{aziz2008modulation} modeled events from on- to offset of sentences
(describing unknown and famous places and faces), and compared activity levels
that were averaged across voxels of ROIs (PPA and fusiform face area, FFA
\citep{kanwisher1997ffa}) defined by a localizer experiment.
% Aziz results
Study results revealed significantly decreased activity in only the left PPA
compared to activity in the FFA, and only for sentences describing famous places
compared to famous faces.
% we: model & whole brain
Here we modeled events from on- to offset of single words and performed a
whole-brain analysis.
% our results
Group results of audio-description's primary contrast yielded spatially concise
clusters in the anterior part of the PPA group overlap.
% concluding statement
Contrary to \citep{aziz2008modulation}, our group results suggest that auditory
spatial information compared to non-spatial information correlates with
bilaterally increased activation in the anterior part of the PPA.

@@

% optimal stimulus type in vision
In the visual domain, landscapes and not pictures of landmarks or buildings are
considered to be the ``optimal'' stimulus type
\citep{epstein2008parahippocampal}.
% we did not choose se\_new and se\_old
We did not include the categories that contained switches from one setting to
another (\texttt{se\_new} and \texttt{se\_old}) which one might assume to
contain the auditory equivalent to the ``optimal'' visual stimulus type in the
primary AD contrast.
% why
The reason was that the categories \texttt{se\_new} and \texttt{se\_old} were
heterogeneous: they rarely contained holistic (but also vague) descriptions of
landscapes (e.g.  ``[Forrest is running through the] jungle'') but mostly
landmarks or buildings, and also non-spatial hints (e.g. ``[Jenny as a]
teenager``).
% further studies
Given that humans can identify the gist of a rich visual scene within the
duration of a single fixation \citep{henderson2003human} further studies might
investigate if vague verbal descriptions of landscapes lead to a different
hemodynamic activity level than descriptions of more concrete parts of a scene
(e.g.``[a] beacon'', [a] ``farmhouse'').

@@

\todo[inline]{how to discuss Bland-Altman-Plots???; dunno if this is meaningful:
z-scores in the anterior part of PPA are higher in AD contrasts than in contrast
of localizer}
%
Based on the fact that the annotation of nouns offered a more precise way to
operationalize the perception of spatial information than the annotation of cuts
and locations, we also hypothesized that analyzing the data from the
audio-description could localize the PPA in individual persons.
%
Hence to test if an auditory narrative could [in principle] substitute a visual
localizer experiment, we examined the results from the audio-description's
primary contrast on an individual level.
% results of dedicated localizer
The visual localizer \citep{sengupta2016extension} revealed bilateral ROIs in 12
of 14 subjects and a unilateral right ROI in two subjects (sub-04, sub-20).
%
Whereas \citep{sengupta2016extension} chose the subjectively ``best fitting''
$z$-threshold in an unthresholded $z$-map from one of three contrasts by visual
judgment, we used a ``one threshold fits all approach`` (Z>3.4, p<.05
cluster-corrected) in one contrast.
% in the current study
The audio-description's primary contrast yielded significant bilateral clusters
in 9 of 14 subjects (of which sub-04 shows only a right-lateralized PPA ROI),
and a unilateral significant cluster in one subject.
% similar: anterior part
Similarly to the group average results, results on an individual level also
revealed that clusters are spatially restricted to the anterior part of
individual PPAs.
% conclusion
Thus, the cluster location and size is not just a by-product of the averaging
process but the same pattern can can reliably be found in individual study
participants.

@@

This raises the question if the different activation patterns can be attributed
to different features inherent in visual compared to verbal stimuli.
% visual scenes: probably submodules
In the domain of visual scene perception, previous studies provide evidence that
the PPA can be divided into functionally subregions that might process different
stimulus features.
% posterior: functionality
The posterior PPA (pPPA) is functionally more responsive than the anterior PPA
(aPPA) to low-level features of scenes or (abstract) objects
\citep{baldassano2013differential, nasr2014thinking,
rajimehr2011parahippocampal}.
% anterior cIPL is defined using the Eickhoff–Zilles PGp probabilistic
% cytoarchitectonic map
The aPPA is functionally more responsive than the pPPA to high-level features of
scenes (e.g. real-word size \citep{park2015parametric}; a scene's abstract
identity/category \citep{marchette2015outside, watson2016patterns}) and objects
(e.g. spatial contextual associations \citep{aminoff2007parahippocampal,
aminoff2013role}).
% connectivity
Moreover, pPPA and aPPA show differences in connectivity profiles.
% posterior
The pPPA is more strongly functionally connected to the occipital visual cortex
than the aPPA \citep{baldassano2013differential, baldassano2016two}.
% including lateral occipital cortex (LOC) and transverse occipital sulcus (TOS)
% anterior
The aPPA is more strongly functionally connected to portions of the default mode
network, including caudal inferior parietal lobe (cIPL), the restrosplenial
complex (RSC), medial PFC and the lateral surface of the anterior temporal lobe
\citep{baldassano2013differential, baldassano2016two}.
% conclusion: subregions
According to the hypothesis by \citep{baldassano2013differential} the PPA
consists of subregions that activate selectively to scenes and ``cooperate to
build a complete representation of a scene''.
% Their distinct connectivity properties do suggest that each may be involved in
% specific aspects of visual and cognitive processing involved in the
% overarching goal of scene understanding \citep{baldassano2013differential}.
% The fact that anterior PPA had a lower sensitivity to our abstract object
% stimuli does not necessarily imply that this region does not use object
% information \citep{baldassano2013differential}. Previous work has shown that
% PPA responds to objects that have spatial associations [Aminoff et al. 2007],
% are space-defining [Mullally and Maguire 2011], and are
% navigationally-relevant [Janzen and Van Turennout 2004]. These types of
% responses require spatial memory and cannot be based purely on visual features
% like object shape. \citep{baldassano2013differential}.
Our results add further evidence that the PPA is not a homogeneous area but can
be divided into functionally subregions.
% pPPA
Our results suggest that the pPPA is more concerned with visual aspects immanent
in pictures or movie shots of landscapes.
% aPPA
The aPPA might be more concerned with non-visual aspects of spatial information
(e.g. spatial contextual information \citep{aminoff2013role,
aminoff2015associative, baumann2016functional}) as immanent/inherent in both
visual and verbal stimuli.

% conclusion
Lastly, our exploratory analysis suggests that a naturalistic auditory stimulus
can [in principle] be applied to localize (the anterior part) of an higher-level
visual area in the majority of subjects.
% one threshold (and contrast) to rule them all
Further studies that aim for individual diagnostics / ROI localization should
nevertheless test more than one contrast to gain the ``best'' individual
ROIs.\todo{..ouff}

@@

% no random clusers
Despite performing a whole brain analysis, we do not find many [random] clusters
in single contrasts or even systematically across contrasts that hamper the
interpretation of results.
% AV: null results regarding attention and eye movements
For example in the AV results, we do not find significant clusters that could be
attributed to cognitive processes like attention or eye movements.

% intro
Apart from the PPA, we find significantly increased activity in the dorsal
precuneus and posterior cingulate region (e.g. retrosplenial cortex; RSC) and
lateral occipital cortex (LOC) across contrasts of both naturalistic stimuli.
% RSC intro
Like the PPA, the RSC and LOC are known to be scene-selective regions (see
\citep{chrastil2018heterogeneity, bettencourt2013role} for reviews).
% conclusion
\todo[inline]{concluding statement}

>>>>

% previous studies
Previous studies reported increased hemodynamic activity in the PPA when
participants were watching static pictures of landscape compared to pictures of
faces or objects (e.g. \citep{epstein1998ppa, epstein1999parahippocampal})
whereas one study using spoken sentences reported mixed results
\citep{aziz2008modulation}.
% current study: similar
In line with previous studies, we operationalized spatial perception via general
linear model (GLM) $t$-contrasts.
% current study: dissimilar
Contrary to previous studies, our current approach is not based on carefully
designed sets of stimuli but based on annotated stimulus features incidentially
occurring in two naturalistic stimuli.
% group AV
On a group level, whole brain-analysis results for the movie stimulus show
signifcantly increased hemodynamic activity spatially overlapping with a
traditionally localized PPA but also extending into earlier visual cortices.
% group AD
Results for the audio-description also reveal significant activation in the PPA
but restricted to its anterior part.
% individual level
On an individual level, results of the audio-description show significant
bilateral clusters in the anterior PPA of nine subjects and significant
unilateral clusters in one subjects.
% meaning: generalization to naturalistic stimuli
Results suggest that increased activation in the PPA during the perception of
static pictures generalizes to the perception of spatial information embedded in
a movie and an exclusively auditory stimulus.
% meaning: aPPA vs. pPPA
Our results provide further evidence that the PPA can be divided into functional
subregions.
% narrative as substitute
An auditory narrative might substitute a visual paradigm to localize
the PPAs anterior part in visually impaired subjects.

@@

% intro
First, we analyzed data from the movie stimulus offering ecologically more valid
but still visual stimulation.
%
An early study \citep{bartels2004mapping} that manually annotated the content of
movie frames (color, faces, language, and human bodies) found that
specialization of functional areas is preserved during movie watching.
%
Contrary to \citep{bartels2004mapping} who annotated the content of frames, we
categorized each frame after a cut independent of its specific content.
\todo{Bartels' clusters are actually pretty big too}
% exploited cinematographic tendencies
We here exploited a cinematographic confound in the stimulus structure (cf.
\ref{rationale}).
% establishing shots vs. e.g. over-the-shoulder shots
Film directors tend to establish the geometry of locations earlier in the movie
and tend to later focus more on persons and objects relevant to the evolved
\citep{brown2012cinematography, mercado2011filmmakers}.
% hypo
Despite our ad hoc approach, we hypothesized that [confounding variables
could/would average out and] our whole-brain analyses would lead to significant
clusters in the PPA.
% results
Group results of the AV stimulus' primary contrast yielded a cluster that span
the group overlap of individual PPA ROIs from anterior to posterior.
% confound of higher perception
The cluster's extension into more posterior, early visual areas can be
attributed to the averaging process that did not control for confounding visual
features [or the attentional focus of study participants].
%
\todo[inline]{I am pretty surprised the clusters are not all over the brain. I
do not know where/how do phrase it but should be mentioned imo}
% solution
Thus to control confounding variables, future studies that aim to use a movie to
localize visual areas should extensively annotate the content of frames (e.g.
using the open-source solution ``Pliers'' for feature extraction from a visual
naturalistic stimulus; \citep{mcnamara2017developing})[, and record the
subject's visual attention via eye-tracking].
%

\todo[inline]{concluding statements regarding exploited cinematographic
confound? e.g.: movies are nice but have idiosyncrasies? so, maybe you should be
aware of these idiosyncrasies (and control for it) in case you wanna use movies
for research?}

\todo[inline]{overall main message / concluding statement of this
``cross-modal-control-movie-thing''??}

@@

% intro
Second, we analyzed data from the audio-description.
% hypothesis
Encouraged by \citep{aziz2008modulation}, we hypothesized that spatial
information embedded in an exclusively auditory stimulus would correlate with
increased hemodynamic activity in the PPA.
% diff to Aziz (2008)
\citep{aziz2008modulation} modeled events from on- to offset of sentences
(describing unknown and famous places and faces), and compared activity levels
that were averaged across voxels of ROIs (PPA and fusiform face area, FFA
\citep{kanwisher1997ffa}) defined by a localizer experiment.
% Aziz results
Study results revealed significantly decreased activity in only the left PPA
compared to activity in the FFA for sentences describing famous places compared
to famous faces.
% we: model & whole brain
Here we modeled events from on- to offset of single words and performed a
whole-brain analysis.
% our results
Group results of audio-description's primary contrast yielded spatially concise
clusters in the anterior part of the PPA group overlap.
% concluding statement
Contrary to \citep{aziz2008modulation}, our group results suggest that auditory
spatial information compared to non-spatial information correlates with
bilaterally increased activation in the anterior part of the PPA.

@@

% optimal stimulus type in vision
In the visual domain, pictures of landscapes and not pictures of landmarks or
buildings are considered to be the ``optimal'' stimulus type
\citep{epstein2008parahippocampal}.
% we did not choose se\_new and se\_old
In the AD's primary contrast, we did not include the categories that contained
switches from one setting to another (\texttt{se\_new} and \texttt{se\_old})
which one might assume to contain the auditory equivalent to pictures of
landscapes.
% why
The reason was that the categories \texttt{se\_new} and \texttt{se\_old} were
heterogeneous: they rarely contained holistic (but also vague) descriptions of
landscapes (e.g.  ``[Forrest is running through the] jungle'') but mostly
landmarks or buildings, and also non-spatial hints (e.g. ``[Jenny as a]
teenager``).
% further studies
Given that humans can identify the gist of a rich visual scene within the
duration of a single fixation \citep{henderson2003human} further studies might
investigate if vague verbal descriptions of landscapes lead to a different
hemodynamic activity level than descriptions of more concrete parts of a scene
(e.g.``[a] beacon'', [a] ``farmhouse'').

@@

\todo[inline]{how to discuss Bland-Altman-Plots???; dunno if this is meaningful:
z-scores in the anterior part of PPA are higher in AD contrasts than in contrast
of localizer}
%
Given that the annotation of nouns offered a more precise way to operationalize
the perception of spatial information than the annotation of cuts and locations,
we also hypothesized that analyzing the data from the audio-description could
localize the PPA in individual persons.
%
We examined the results from the audio-description's primary contrast on an
individual level to test if an auditory narrative could [in principle]
substitute a visual localizer experiment.
% results of dedicated localizer
The visual localizer \citep{sengupta2016extension} revealed bilateral ROIs in 12
of 14 subjects and a unilateral right ROI in two subjects (sub-04, sub-20).
%
Whereas \citep{sengupta2016extension} chose the subjectively ``best fitting''
$z$-threshold in an unthresholded $z$-map from one of three contrasts by visual
judgment, we used one contrast and a ``one threshold fits all approach`` (Z>3.4,
p<.05 cluster-corrected).
% in the current study
The audio-description's primary contrast yielded significant bilateral clusters
in 9 of 14 subjects (of which sub-04 shows only a right-lateralized PPA ROI),
and a unilateral significant cluster in one subject.
% similar: anterior part
Similarly to the group average results, results on an individual level also
revealed that clusters are spatially restricted to the anterior part of
individual PPAs.
% conclusion
Thus, the cluster location and size found on the group level is not just a
by-product of the averaging process but the same pattern can can reliably be
found in individual study participants.
% concluding statement
our exploratory analysis suggests that a naturalistic auditory stimulus
can [in principle] be applied to localize (the anterior part) of an higher-level
visual area in the majority of subjects.
% one threshold (and contrast) to rule them all
Further studies that aim for individual diagnostics / ROI localization should
nevertheless test more than one contrast to gain the ``best'' individual
ROIs.\todo{..ouff}

@@

% question
The fact that clusters from the auditory stimulus are spatially restricted to
the anterior part of clusters from the visual paradigms raises the question if
the revealed activation pattern could be attributed to be attributed to
different features inherent in visual compared to verbal stimuli.\todo{ouff}
% visual scenes: probably submodules
In the domain of visual scene perception, previous studies provide evidence that
the PPA can be divided into functionally subregions that might process different
stimulus features.
% posterior: functionality
The posterior PPA (pPPA) is functionally more responsive than the anterior PPA
(aPPA) to low-level features of scenes or (abstract) objects
\citep{baldassano2013differential, nasr2014thinking,
rajimehr2011parahippocampal}.
% anterior cIPL is defined using the Eickhoff–Zilles PGp probabilistic
% cytoarchitectonic map
The aPPA is functionally more responsive than the pPPA to high-level features of
scenes (e.g. real-word size \citep{park2015parametric}; a scene's abstract
identity/category \citep{marchette2015outside, watson2016patterns}) and objects
(e.g. spatial contextual associations \citep{aminoff2007parahippocampal,
aminoff2013role}).
% connectivity
Moreover, pPPA and aPPA show differences in connectivity profiles.
% posterior
The pPPA is more strongly functionally connected to the occipital visual cortex
than the aPPA \citep{baldassano2013differential, baldassano2016two}.
% including lateral occipital cortex (LOC) and transverse occipital sulcus (TOS)
% anterior
The aPPA is more strongly functionally connected to portions of the default mode
network, including caudal inferior parietal lobe (cIPL), the restrosplenial
complex (RSC), medial PFC and the lateral surface of the anterior temporal lobe
\citep{baldassano2013differential, baldassano2016two}.
% conclusion: subregions
According to the hypothesis by \citep{baldassano2013differential} the PPA
consists of subregions that activate selectively to scenes and ``cooperate to
build a complete representation of a scene''.
% Their distinct connectivity properties do suggest that each may be involved in
% specific aspects of visual and cognitive processing involved in the
% overarching goal of scene understanding \citep{baldassano2013differential}.
% The fact that anterior PPA had a lower sensitivity to our abstract object
% stimuli does not necessarily imply that this region does not use object
% information \citep{baldassano2013differential}. Previous work has shown that
% PPA responds to objects that have spatial associations [Aminoff et al. 2007],
% are space-defining [Mullally and Maguire 2011], and are
% navigationally-relevant [Janzen and Van Turennout 2004]. These types of
% responses require spatial memory and cannot be based purely on visual features
% like object shape. \citep{baldassano2013differential}.
Our results add further evidence that the PPA is not a homogeneous area but can
be divided into functionally subregions.
% pPPA
Our results suggest that the pPPA is more concerned with visual aspects immanent
in pictures or movie shots of landscapes.
% aPPA
The aPPA might be more concerned with non-visual aspects of spatial information
(e.g. spatial contextual information \citep{aminoff2013role,
aminoff2015associative, baumann2016functional}) as immanent/inherent in both
visual and verbal stimuli.

@@

% no random clusters
Despite performing a whole brain analysis, we do not find clusters in the
primary contrast or across contrasts (s. supplemental material) that hamper the
interpretation of results.
% intro
Apart from the PPA, we find significantly increased activity in the dorsal
precuneus and posterior cingulate region (e.g. retrosplenial complex; RSC) and
lateral occipital cortex (LOC) across contrasts of both naturalistic stimuli.
% RSC intro
Like the PPA, the RSC and LOC repeatedly showed increased hemodynamic activity
in studies investigating visual spatial perception or navigation (see
\citep{chrastil2018heterogeneity, bettencourt2013role} for reviews).

\todo[inline]{AD clusters in LOC are more medial in AD stimulus compared to AV;
whereas clusters in PPA and RSC of both stimuli overlap}
% conclusion
\todo[inline]{concluding statement??}

This suggest that model-driven methods based on stimulus annotations are
suitable to reveal spatially concise functional areas and networks.
%
Nevertheless, ``cleanliness''/``neatness``  of results (i.e. the signal-cluster
to noise-cluster ration) depend on the extent to which the multi-dimensionality
of the naturalistic stimulus is annotated, modeled, statistically controlled and
/ or how many events the stimulus provides per assumed perceptual process for
the averaging purpose

\section{summary}

\todo[inline]{join summary \& conclusion}

% summary
In summary, our approach to operationalize spatial perception based on stimulus
annotations of two naturalistic stimuli is the shit and provide further evidence
that model-driven analyses can be used to analyze data from multi-dimensional
stimuli
%
and before I forget: this study shows that already existing, publicly
available data can be re-used for independent research questions that provide
further evidence for existing theories/hypotheses and advance the field by
generating new findings \& hypothesis.\todo{if there are any parts that can be
rephrased as explicit hypothesis: do it!}
